---
title: "Code for Using Data in MySQL"
output: github_document
---

##### _Loading Required Packages_  
```{r loading needed packages, message = FALSE}
library(RMySQL)
library(tidyverse)
library(keyring)
```

The `keyring` package is needed in order to maintain the security and privacy of our login information. This is crucial because we do not want to share login information when we distribute our code to others. We generate a key for our password and use it by calling it with the `key_get()` function in the `keyring` package. Coding this key is done directly in the console and not included in the final code.  

***  

##### _Connecting to the MySQL Schema "Sakila"_  
```{r connecting to local MySQL database}
sakila_db = dbConnect(MySQL(),
            user = "root",
            password = key_get("root_pw"),
            dbname = 'sakila',
            host = 'localhost')
```

The `Sakila` schema is a sample database that contains information on movie rental customers, as well as information on the movies themselves. You can find both the SQL schema documentation and zip file for download [here](https://dev.mysql.com/doc/sakila/en/) and [here](https://downloads.mysql.com/docs/sakila-db.zip), respectively.

